Important Factors in Predicting a Stroke

### 2021-05-13 draft.Rmd
more exploratory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyverse)
library(caret)
library(DT)
```

@@

-initial summary statistics

@@

-look at correlations between variables
-heart disease and hypertension?

>>>>

```{r setup, include=FALSE}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyverse)
library(caret)
library(DT)
```

@@

@@

#### Summary Statistics

Initial summary statistics, after re-classing some of the variables:

```{r, echo = FALSE}
# changing variable types for analysis and replacing missing values
stroke$bmi <- gsub("N/A", NA, stroke$bmi)
stroke$smoking_status <- gsub("Unknown", NA, stroke$smoking_status)
stroke$gender <- as.factor(stroke$gender)
stroke$bmi <- as.numeric(stroke$bmi)
stroke$age <- as.numeric(stroke$age)
stroke$avg_glucose_level <- as.numeric(stroke$avg_glucose_level)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$work_type <- as.factor(stroke$work_type)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)

summary(stroke)
```

The dataset is pretty evenly split among gender.

```{r, echo = FALSE}
stroke <- as_tibble(stroke)
barplot(table(stroke$gender))
```

The two variables were are most interested in are not very balanced. On the left is a histogram for hypertension, and on the right is one for heart disease.As mentioned before, only 9.75% of the patients having hypertension and 5.40% of them having heart disease

```{r, echo = FALSE, figures-side, fig.show = "hold", out.width="50%"}
stroke <- as_tibble(stroke)
barplot(table(stroke$hypertension))
barplot(table(stroke$heart_disease))
```

Smoking and BMI could also be of interest.

```{r, echo = FALSE, fig.show = "hold", out.width="50%"}
stroke <- as_tibble(stroke)
barplot(table(stroke$smoking_status))
barplot(table(stroke$bmi))
```

The average BMI is 28.89, which is classified as overweight. Being overweight or obese increases your risk for heart disease and stroke (but then again BMI isn't a completely accurate measurement). Based on the information from the CDC, we would expect patients who smoke or even formerly smoked or with higher BMIs would have a higher incidence of stroke.

#### Visualizations

```{r, echo = FALSE}
# plotting the distribution of bmi based on smoking status

# plotting distribution of bmi based on hypertension

#plotting sitribution of bmi based on heart disease

#plotting distribution of bmi based on incidence of stroke?

# strip plots?
```
